Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/instructions.tex
upd
add filenames

If we left Routine initialization to here we can leave method
initialization. \#\# Method initialization

@@

call Config ObjectNew

\#Po00ol = \#Outcome

Set self and super

@@

If the routine is invoked as a function, \#lsFunction.\#NewLevel shall
be set to `1', otherwise to `0O'; this affects the processing of a
subsequent RETURN instruction.

\#AllowProcedure.\#NewLevel = `1'

Many of the initial values for a new invocation are inherited from the
caller's values. \#Digits.\#NewLevel = \#Digits.\#Level

\#Form.\#NewLevel = \#Form.\#Level

\#Fuzz.\#NewLevel = \#Fuzz.\#Level

\#StartTime.\#NewLevel = \#StartTime.\#Level

\#Tracing.\#NewLevel = \#Tracing.\#Level \#Interactive.\#NewLevel =
\#Interactive.\#Level

call EnvAssign ACTIVE, \#NewLevel, ACTIVE, \#Level call EnvAssign
ALTERNATE, \#NewLevel, ALTERNATE, \#Level

do t=1 to 7 Condition = word(`SYNTAX HALT ERROR FAILURE NOTREADY NOVALUE
LOSTDIGITS',t) \#Enabling.Condition.\#NewLevel =
\#Enabling.Condition.\#Level

\#Instruction.Condition.\#NewLevel = \#Instruction.Condition.\#Level
\#TrapName.Condition.\#NewLevel = \#TrapName.Condition.\#Level
\#EventLevel.Condition.\#NewLevel = \#EventLevel.Condition.\#Level end t

If this invocation is not caused by a condition occurring, see nnn, the
state variables for the CONDITION

built-in function are copied. \#Condition.\#NewLevel =
\#Condition.\#Level \#ConditionDescription.\#NewLevel =
\#ConditionDescription.\#Level \#ConditionExtra.\#NewLevel =
\#ConditionExtra.\#Level \#ConditionInstruction.\#NewLevel =
\#ConditionInstruction.\#Level Execution of the initialized routine
continues at the new level of invocation. \#Level = \#NewLevel
\#NewLevel = \#Level + 1

@@

The clause is traced before execution: if pos(\#Tracing.\#Level, `AIR')
\textgreater{} 0 then call \#TraceSource

@@

\#ClauseTime.\#Level = '\,' The state variable \#LineNumber is set to
the line number of the clause, see nnn. A clause other than a null
clause or label or procedure instruction sets:

\#AllowProcedure.\#Level = `0' /* See message 17.1 */

@@

if \#InhibitTrace \textgreater{} 0 then \#InhibitTrace = \#InhibitTrace
- 1 Polling for a HALT condition occurs:

\#Response = Config Halt Query ()

if \#0utcome == `Yes' then do call Config Halt Reset

call \#Raise `HALT', substr(\#Response,2) /* May return */ end

At the end of each clause there is a check for conditions which occurred
and were delayed. It is acted on

if this is the clause in which the condition arose. do t=1 to 4
\#Condition=WORD(`HALT FAILURE ERROR NOTREADY',t) /* HALT can be
established during HALT handling. */ do while
\#PendingNow.\#Condition.\#Level \#PendingNow.\#Condition.\#Level = `0'
call \#Raise end end

Interactive tracing may be turned on via the configuration. Only a
change in the setting is significant. call Config Trace Query

if \#AtPause = 0 \& \#Outcome == `Yes' \& \#Trace QueryPrior == `No'
then do /* External request for Trace `?R!' */ \#Interactive.\#Level =
`1' \#Tracing.\#Level = `R' end

\#TraceQueryPrior = \#Outcome

Tracing just not the same with NetRexx.

@@

At a pause point: if \#AtPause = 0 \& \#Interactive.\#Level \&
\#InhibitTrace = 0 then do if \#InhibitPauses \textgreater{} 0 then
\#InhibitPauses = \#InhibitPauses-1 else do \#TraceInstruction = `0' do
forever

call Config Trace Query

if \#Outcome == `No' \& \#Trace QueryPrior == `Yes' then do /* External
request to stop tracing. */ \#Trace\_QueryPrior=\#Outcome
\#Interactive.\#Level = `0' \#Tracing.\#Level = `N' leave end

if \#Outcome == `Yes' \& \#Trace QueryPrior == `No' then do /* External
request for Trace `?R!' */ \#Trace QueryPrior = \#Outcome
\#Interactive.\#Level = `1' \#Tracing.\#Level = `R' leave end

if \#Interactive.\#Level \textbar{} \#TraceInstruction then leave

/* Accept input for immediate execution. */

call Config Trace Input

if length(\#0utcome) = 0 \textbar{} \#0Outcome == `=' then leave
\#AtPause = \#Level

interpret \#Outcome

\#AtPause = 0 end /* forever loop \emph{/ if \#Outcome == `=' then call
\#Retry /} With no return */ end end

@@

I/O can be redirected when submitting commands to an external
environment. The submitted command's input stream can be taken from an
existing stream or from a set of compound variables with a common stem.
In the latter case (that is, when a stem is specified as the source for
the commands input stream) whole number tails are used to order input
for presentation to the submitted command. Stem.0 must contain a whole
number indicating the number of compound variables to be presented, and
stem. 1 through stem.n (where n=stem.0) are the compound variables to be
presented to the submitted command.

@@

The redirections specified on the ADDRESS instruction may not be
possible. If the configuration is aware that the command processor named
does not perform I/O in a manner compatible with the request, the value
of \#Env\_Type. may be set to `UNUSED' as an alternative to `STEM' and
`STREAM' where those values are assigned in the following code.

In the following code the particular use of \#Contains(address,
expression) refers to an expression immediately contained in the
address.

Addrinstr: /* If ADDRESS keyword alone, environments are swapped.
\emph{/ if \#Contains (address, taken constant), \& \#Contains
(address,valueexp), \& \#Contains (address, `WITH') then do call
EnvAssign TRANSIENT, \#Level, ACTIVE, \#Level call EnvAssign ACTIVE,
\#Level, ALTERNATE, \#Level call EnvAssign ALTERNATE, \#Level,
TRANSIENT, \#Level return end /} The environment name will be explicitly
specified. */ if \#Contains(address,taken constant) then Name =
\#Instance(address, taken \_ constant) else Name = \#Evaluate(valueexp,
expression) if length(Name) \textgreater{} \#LimitEnvironmentName then
call \#Raise `SYNTAX', 29.1, Name

if \#Contains(address,expression) then do /* The command is evaluated
(but not issued) at this point. */ Command = \#Evaluate (address,
expression)

if \#Tracing.\#Level == `C' \textbar{} \#Tracing.\#Level == `A' then do
call \#Trace '\textgreater\textgreater\textgreater! end

end

call AddressSetup /* Note what is specified on the ADDRESS instruction.
\emph{/ /} If there is no command, the persistent environment is being
set. */ if \#Contains(address,expression) then do

call EnvAssign ACTIVE, \#Level, TRANSIENT, \#Level

return

end

call CommandIssue Command /* See nnn \emph{/ return /} From Addrinstr */

AddressSetup: /* Note what is specified on the ADDRESS instruction, into
the TRANSIENT environment. \emph{/ EnvTail = `TRANSIENT.'\#Level /}
Initialize with defaults. */ \#Env\_Name.EnvTail = '\,'

\#Env\_ Type.I.EnvTail = `NORMAL' \#Env\_ Type.O.EnvTail = `NORMAL'
\#Env\_ Type.E.EnvTail = `NORMAL'

\#Env\_Resource.I.EnvTail = '\,' \#Env\_Resource.O.EnvTail = `!
\#Env\_Resource.E.EnvTail ='' /* APPEND / REPLACE does not apply to
input. */

\#Env\_Position.I.EnvTail = `INPUT' \#Env\_Position.O.EnvTail =
`REPLACE' \#Env\_Position.E.EnvTail = `REPLACE'

/* If anything follows ADDRESS, it will include the command processor
name.*/ \#Env\_Name.EnvTail = Name

/* Connections may be explicitly specified. */ if \#Contains (address,
connection) then do

if \#Contains(connection,input) then do /* input redirection */ if
\#Contains (resourcei, `STREAM') then do \#Env\_Type.I.EnvTail =
`STREAM' \#Env\_Resource.1.EnvTail=\#Evaluate(resourcei, VAR\_SYMBOL)
end if \#Contains (resourcei, `STEM') then do \#Env\_Type.I.EnvTail =
`STEM'

Temp=\#Instance (resourcei,VAR\_SYMBOL) if \#Parses(Temp, stem /* See
nnn \emph{/) then call \#Raise `SYNTAX', 53.3, Temp
\#Env\_Resource.I.EnvTail=Temp end end /} Input */

if \#Contains(connection,output) then /* output redirection */ call
NoteTarget O

if \#Contains(connection,error) then /* error redirection \emph{/ /} The
prose on the description of \#Contains specifies that the relevant
resourceo is used in NoteTarget. */ call NoteTarget E

end /* Connection */

return /* from AddressSetup */

NoteTarget:

/* Note the characteristics of an output resource. */

arg Which /* O or E */

if \#Contains (resourceo, `STREAM') then do
\#Env\_Type.Which.EnvTail=`STREAM'
\#Env\_Resource.Which.EnvTail=\#Evaluate(resourceo, VAR\_SYMBOL) end

if \#Contains(resourceo,`STEM') then do \#Env\_Type.Which.EnvTail=`STEM'
Temp=\#Instance (resourceo, VAR\_SYMBOL) if \#Parses(Temp, stem /* See
nnn */) then

call \#Raise `SYNTAX', 53.3, Temp

\#Env\_Resource.Which.EnvTail=Temp end

if \#Contains (resourceo,append) then
\#Env\_Position.Which.EnvTail=`APPEND' return /* From NoteTarget */

EnvAssign: /* Copy the values that name an environment and describe its
redirections. */ arg Lhs, LhsLevel, Rhs, RhsLevel
\#Env\_Name.Lhs.LhsLevel = \#Env\_Name.Rhs.RhsLevel \#Env\_
Type.I.Lhs.LhsLevel = \#Env\_Type.I.Rhs.RhsLevel \#Env\_
Resource.I.Lhs.LhsLevel = \#Env\_Resource.I.Rhs.RhsLevel
\#Env\_Position.I.Lhs.LhsLevel = \#Env\_Position.I.Rhs.RhsLevel \#Env\_
Type.O.Lhs.LhsLevel = \#Env\_Type.O.Rhs.RhsLevel
\#Env\_Resource.O.Lhs.LhsLevel = \#Env\_Resource.O.Rhs.RhsLevel
\#Env\_Position.O.Lhs.LhsLevel = \#Env\_Position.O.Rhs.RhsLevel \#Env\_
Type.E.Lhs.LhsLevel = \#Env\_Type.E.Rhs.RhsLevel
\#Env\_Resource.E.Lhs.LhsLevel \#Env\_Resource.E.Rhs.RhsLevel
\#Env\_Position.E.Lhs.LhsLevel \#Env\_Position.E.Rhs.RhsLevel return

@@

The ARG instruction is a shorter form of the equivalent instruction:
PARSE UPPER ARG template list

@@

Assignment can occur as the result of executing a clause containing an
assignment (see nnn and nnn), or as a result of executing the VALUE
built-in function, or as part of the execution of a PARSE instruction.
Assignment involves an expression and a VAR\_SYMBOL. The value of the
expression is determined; see nnn.

If the VAR\_SYMBOL does not contain a period, or contains only one
period as its last character, the

value is associated with the VAR\_SYMBOL: call Var Set \#Pool,VAR
SYMBOL, `0',Value

Otherwise, a name is derived, see nnn. The value is associated with the
derived name: call Var Set \#Pool,Derived Name,`1',Value

@@

The CALL instruction is used to invoke a routine, or is used to control
trapping of conditions. lf a vrefis specified that value is the name of
the routine to invoke:

if \#Contains (call, vref) then Name = \#Evaluate(vref, var\_symbol)

If a taken\_constant is specified, that name is used. if \#Contains
(call, taken constant) then Name = \#Instance(call, taken constant)

The name is used to invoke a routine, see nnn. If that routine does not
return a result the RESULT and

-RESULT variables become uninitialized: call Var Drop \#Pool, `RESULT',
`0! call Var Drop \#ReservedPool,'.RESULT', `0'

@@

If the routine returns a result and the trace setting is `R' or `I' then
a trace with that result and a tag
'\textgreater\textgreater\textgreater'' shall be produced, associated
with the call instruction.

If a callon\_spec is specified: If \#Contains(call,callon spec) then do
Condition = \#Instance(callon\_spec,callable condition)

\#Instruction.Condition.\#Level = `CALL' If \#Contains(callon spec,
`OFF') then \#Enabling.Condition.\#Level = `OFF' else
\#Enabling.Condition.\#Level = `ON'

/* Note whether NAME supplied. */ If Contains (callon spec,taken
constant) then Name = \#Instance (callable condition, taken\_constant)

else

Name = Condition \#TrapName.Condition.\#Level = Name end

@@

For a definition of the syntax of a command, see nnn. A command that is
not part of an ADDRESS instruction is processed in the ACTIVE
environment.

Command = \#Evaluate(command, expression)

if \#Tracing.\#Level == `C' \textbar{} \#Tracing.\#Level == `A' then
call \#Trace `\textgreater\textgreater\textgreater!'

call EnvAssign TRANSIENT, \#Level, ACTIVE, \#Level

call CommandiIssue Command

Commandlssue is also used to describe the ADDRESS instruction:

CommandIssue: parse arg Cmd /* Issues the command, requested environment
is TRANSIENT \emph{/ /} This description does not require the command
processor to understand stems, so it uses an altered environment. */
call EnvAssign PASSED, \#Level, TRANSIENT, \#Level EnvTail =
`TRANSIENT.'\#Level

/* Note the command input. \emph{/ if \#Env\_Type.I.EnvTail = `STEM'
then do /} Check reasonableness of the stem. \emph{/ Stem =
\#Env\_Resource.I.EnvTail Lines = value(Stem'0') if \datatype(Lines,`W')
then call \#Raise `SYNTAX',54.1,Stem'0', Lines if Lines\textless0 then
call \#Raise `SYNTAX',54.1,Stem'0', Lines /} Use a stream for the stem
*/ \#Env\_ Type.I.PASSED.\#Level = `STREAM' call Config Stream Unique
InputStream = \#Outcome \#Env\_Resource.1I.PASSED.\#Level = InputStream
call charout InputStream , vl do j = 1 to Lines call lineout
InputStream, value(Stem \textbar\textbar{} j) end j call lineout
InputStream end

/* Note the command output. */

if \#Env\_Type.O.EnvTail = `STEM' then do Stem =
\#Env\_Resource.O.EnvTail if \#Env\_Position.O.EnvTail == `APPEND' then
do

/* Check that Stem.0 will accept incrementing. \emph{/ Lines=value
(Stem'0'); if \datatype(Lines,`W') then call \#Raise
`SYNTAX',54.1,Stem'0', Lines if Lines\textless0 then call \#Raise
`SYNTAX',54.1,Stem'0', Lines end else call value Stem'0',O /} Use a
stream for the stem */ \#Env\_Type.O.PASSED.\#Level = `STREAM' call
Config Stream Unique \#Env\_Resource.O.PASSED.\#Level = \#Outcome end

/* Note the command error stream. */

if \#Env\_Type.E.EnvTail = `STEM' then do Stem =
\#Env\_Resource.E.EnvTail if \#Env\_Position.E.EnvTail == `APPEND' then
do

/* Check that Stem.0 will accept incrementing. */ Lines=value (Stem'0');
if \datatype(Lines,`W') then call \#Raise `SYNTAX',54.1,Stem'0', Lines
if Lines\textless0 then call \#Raise `SYNTAX',54.1,Stem'0', Lines

end else call value Stem'0',0O /* Use a stream for the stem */ \#Env\_
Type.E.PASSED.\#Level = `STREAM' call Config Stream Unique
\#Env\_Resource.E.PASSED.\#Level = \#Outcome end

\#API Enabled = `1'

call Var\_Reset \#Pool

/* Specifying PASSED here implies all the

components of that environment. */

\#Response = Config Command (PASSED, Cmd) \#Indicator = left
(\#Response,1) Description = substr (\#Response, 2)

\#API Enabled = `0'

/* Recognize success and failure. */ if \#AtPause = 0 then do

call value `RC', \#RC

call var Set 0, `.RC', 0, \#RC end

select when \#Indicator==`N' then Temp=0

when \#Indicator==`F' then Temp=-1 /* Failure \emph{/ when
\#Indicator==`E' then Temp=1 /} Error \emph{/ end call Var Set 0, `.RS',
0, Temp /} Process the output \emph{/ if \#Env\_Type.O.EnvTail=`STEM'
then do /} get output into stem. \emph{/ Stem =
\#Env\_Resource.0O.EnvTail OutputStream =
\#Env\_Resource.0O.PASSED.\#Level do while lines (OutputStream)
\textgreater{} 0 call value Stem'0O',value(Stem'0')4+1 call value
Stem\textbar{} \textbar value(Stem'0'),linein (OutputStream) end end /}
Stemmed Output \emph{/ if \#Env\_Type.E.EnvTail=`STEM' then do /} get
error output into stem. */ Stem = \#Env\_Resource.E.EnvTail OutputStream
= \#Env\_Resource.E.PASSED.\#Level do while lines (OutputStream)
\textgreater{} 0 call value Stem'0O',value(Stem'0')4+1 call value
Stem\textbar{} \textbar value(Stem'0'),linein (OutputStream)

end end /* Stemmed Error output */ if \#Indicator == `N' \&
pos(\#Tracing.\#Level, `CAIR') \textgreater{} 0 then call \#Trace `+++'
if (\#Indicator == `N' \& \#Tracing.\#Level==`E'), \textbar{}
(\#Indicator==`F' \& (\#Tracing.\#Level==`F' \textbar{}
\#Tracing.\#Level==`N')) then do

call \#Trace `\textgreater\textgreater\textgreater!' call \#Trace `+++!'

end \#Condition=`FAILURE' if \#Indicator=`F' \&
\#Enabling.\#Condition.\#Level == `OFF' then call \#Raise `FAILURE' ,
Cmd else if \#Indicator=`E' \textbar{} \#Indicator=`F' then call \#Raise
`ERROR', Cmd

return /* From CommandIssue */

@@

The DO instructions is used to group instructions together and
optionally to execute them repeatedly. Executing a do\_simple has the
same effect as executing a nop, except in its trace output. Executing
the do\_ending associated with a do\_simple has the same effect as
executing a nop, except in its trace output.

A do\_instruction that does not contain a do\_simple is equivalent,
except for trace output, to a sequence of instructions in the following
order.

\#Loop = \#Loop+1 \#Iterate.\#Loop = \#Clause (IterateLabel)
\#Once.\#Loop = \#Clause (OnceLabel) \#Leave.\#Loop = \#Clause
(LeaveLabel) if \#Contains (do specification,assignment) then
\#Identity.\#Loop = \#Instance(assignment, VAR SYMBOL) if \#Contains (do
specification, repexpr) then if \datatype(repexpr,`W') then call \#Raise
`SYNTAX', 26.2,repexpr else do \#Repeat.\#Loop = repexpr+0 if
\#Repeat.\#Loop\textless0 then call \#Raise
`SYNTAX',26.2,\#Repeat.\#Loop end if \#Contains (do
specification,assignment) then do \#StartValue.\#Loop = \#Evaluate
(assignment, expression) if datatype (\#StartValue.\#Loop) == `NUM' then
call \#Raise `SYNTAX', 41.6, \#StartValue.\#Loop \#StartValue.\#Loop =
\#StartValue.\#Loop + 0 if \#Contains (do specification,byexpr) then
\#By.\#Loop = 1 end

The following three assignments are made in the order in which `TO'``,
`BY' and `FOR' appear in docount; see nnn.

if \#Contains (do specification, toexpr) then do if datatype(toexpr) ==
`NUM' then call \#Raise `SYNTAX', 41.4, toexpr \#To.\#LOop = toexpr+0 if
\#Contains (do specification, byexpr) then do if datatype (byexpr)
==`NUM' then call \#Raise `SYNTAX', 41.5, byexpr \#By.\#Loop = byexpr+0
if \#Contains (do specification, forexpr) then do if \datatype(forexpr,
`W') then call \#Raise `SYNTAX', 26.3, forexpr \#For.\#Loop = forexpr+0
if \#For.\#Loop \textless0 then call \#Raise `SYNTAX', 26.3,
\#For.\#Loop end if \#Contains (do specification,assignment) then do
call value \#Identity.\#Loop, \#StartValue.\#Loop end if \#Contains (do
specification, `OVER') then do Value = \#Evaluate(dorep, expression)
\#OverArray.\#Loop = Value \textasciitilde{} makearray

\#Repeat.\#Loop = \#OverArray\textasciitilde items /* Count this
downwards as if repexpr. */ \#Iidentity.\#Loop = \#Instance(dorep, VAR
SYMBOL) end

call \#Goto \#Once.\#Loop /* to OnceLabel */

IterateLabel:

if \#Contains (do specification, untilexpr) then do Value =
\#Evaluate(untilexp, expression)

if Value == `1' then leave if Value == `0' then call \#Raise `SYNTAX',
34.4, Value end

if \#Contains (do specification, assignment) then do t = value
(\#Identity. \#Loop)

if \#Indicator == `D' then call \#Raise `NOVALUE', \#Identity.\#Loop
call value \#Identity.\#Loop, t + \#By.\#Loop end

OnceLabel:

if \#Contains (do specification, toexpr) then do if
\#By.\#Loop\textgreater=0 then do if value(\#Identity.\#Loop)
\textgreater{} \#To.\#Loop then leave end else do if
value(\#Identity.\#Loop) \textless{} \#To.\#Loop then leave end end

if \#Contains(dorep, repexpr) \textbar{} \#Contains(dorep, `OVER') then
do if \#Repeat.\#Loop = 0 then leave \#Repeat.\#Loop = \#Repeat.\#Loop-1
if \#Contains(dorep, `OVER') then call value \#Identity.\#Loop,
\#OverArray{[}\#OverArray\textasciitilde items - \#Repeat.\#Loop{]} end
if \#Contains (do specification, forexpr) then do if \#For.\#Loop = 0
then leave \#For.\#Loop = \#For.\#Loop - 1 end if \#Contains (do
specification, whileexpr) then do Value = \#Evaluate(whileexp,
expression)

if Value == `0' then leave if Value == `1' then call \#Raise `SYNTAX',
34.3, Value end \#Execute (do instruction, instruction list) TraceOfEnd:
call \#Goto \#Iterate.\#Loop /* to IterateLabel */ LeaveLabel:

\#Loop = \#Loop - 1

@@

When clauses are being traced by \#TraceSource, due to
pos(\#Tracing.\#Level, `AIR') \textgreater{} 0, the DO instruction shall
be traced when it is encountered and again each time the IterateLabel
(see nnn) is encountered. The END instruction shall be traced when the
TraceOfEnd label is encountered.

When expressions or intermediates are being traced they shall be traced
in the order specified by nnn. Hence, in the absence of conditions
arising, those executed prior to the first execution of OnceLabel shall
be shown once per execution of the DO instruction; others shall be shown
depending on the outcome of the tests.

The code in the DO description: t = value (\#Identity. \#Loop) if
\#Indicator == `D' then call \#Raise `NOVALUE', \#Identity.\#Loop call
value \#Identity.\#Loop, t + \#By.\#Loop

represents updating the control variable of the loop. That assignment is
subject to tracing, and other expressions involving state variables are
not. When tracing intermediates, the BY value will have a tag of
`\textgreater+\textgreater{}',

@@

The words of the variable\_list are processed from left to right.

A word which is a VAR\_SYMBOL, not contained in parentheses, specifies a
variable to be dropped. If VAR\_SYMBOL does not contain a period, or has
only a single period as its last character, the variable

associated with VAR\_SYMBOL by the variable pool is dropped:

\#Response = Var Drop (\#Pool,VAR\_ SYMBOL, `0')

@@

\#Response = Var Drop (\#Pool,VAR SYMBOL, `1')

If the word of the variable\_list is a VAR\_SYMBOL enclosed in
parentheses then the value of the

VAR\_SYMBOL is processed. The value is considered in uppercase: \#Value
= Config Upper (\#Value)

@@

If the word does not have the syntax of VAR\_SYMBOL a condition is
raised: call \#Raise `SYNTAX', 20.1, word

Otherwise the VAR\_SYMBOL indicated by the word is dropped, as if that
VAR\_SYMBOL were a word of the variable\_list.

@@

Any expression is evaluated:

if \#Contains(exit, expression) then Value = \#Evaluate(exit,
expression) \#Level = 1

\#Pool = \#Pooll

The opportunity is provided for a final trap. \#API Enabled = `1' call
Var\_Reset \#Pool call Config Termination \#API Enabled = `0'

The processing of the program is complete. See nnn for what API Start
returns as the result.

@@

We need a check that this starts method; similarities with PROCEDURE.

@@

It is used at the start of a method, after method initialization, to
make variables in the receiver's pool

accessible: if \#AllowExpose then call \#Raise `SYNTAX', 17.2

The words of the variable\_list are processed from left to right.

A word which is a VAR\_SYMBOL, not contained in parentheses, specifies a
variable to be made accessible. If VAR\_SYMBOL does not contain a
period, or has only a single period as its last character, the variable
associated with VAR\_SYMBOL by the variable pool (as a non-tailed name)
is given the

>>>>

\emph{If we left Routine initialization to here we can leave method
initialization.}

\section{Method initialization}\label{method-initialization}

@@

\lstinputlisting[language=rexx,label=localvariablepoolinit.rexx,caption=localvariablepoolinit.rexx]{localvariablepoolinit.rexx}

\emph{Set self and super}

@@

If the routine is invoked as a function,
\texttt{\#IsFunction.\#NewLevel} shall be set to
\texttt{\textquotesingle{}1\textquotesingle{}}, otherwise to
\texttt{\textquotesingle{}0\textquotesingle{}}; this affects the
processing of a subsequent RETURN instruction.

\lstinputlisting[language=rexx,label=routineinit.rexx,caption=routineinit.rexx]{routineinit.rexx}

Many of the initial values for a new invocation are inherited from the
caller's values.

\lstinputlisting[language=rexx,label=callersvalues.rexx,caption=callersvalues.rexx]{callersvalues.rexx}

If this invocation is not caused by a condition occurring, see nnn, the
state variables for the CONDITION built-in function are copied.

\lstinputlisting[language=rexx,label=conditionlevel.rexx,caption=conditionlevel.rexx]{conditionlevel.rexx}

Execution of the initialized routine continues at the new level of
invocation.

\lstinputlisting[language=rexx,label=newlevel.rexx,caption=newlevel.rexx]{newlevel.rexx}

@@

The clause is traced before execution:

\lstinputlisting[language=rexx,label=traceclauseinit.rexx,caption=traceclauseinit.rexx]{traceclauseinit.rexx}

@@

\lstinputlisting[language=rexx,label=clauseretaindatetime.rexx,caption=clauseretaindatetime.rexx]{clauseretaindatetime.rexx}

The state variable \texttt{\#LineNumber} is set to the line number of
the clause, see nnn.

A clause other than a null clause or label or procedure instruction
sets:

\lstinputlisting[language=rexx,label=allowprocedure.rexx,caption=allowprocedure.rexx]{allowprocedure.rexx}

@@

\lstinputlisting[language=rexx,label=inihibittrace.rexx,caption=inihibittrace.rexx]{inihibittrace.rexx}

Polling for a HALT condition occurs:

\lstinputlisting[language=rexx,label=haltquery.rexx,caption=haltquery.rexx]{haltquery.rexx}

At the end of each clause there is a check for conditions which occurred
and were delayed. It is acted on if this is the clause in which the
condition arose.

\lstinputlisting[language=rexx,label=endclauseconditions.rexx,caption=endclauseconditions.rexx]{endclauseconditions.rexx}
be turned on via the configuration. Only a change in the setting is
significant.

\lstinputlisting[language=rexx,label=configtracequery.rexx,caption=configtracequery.rexx]{configtracequery.rexx}

\emph{Tracing just not the same with NetRexx.}

@@

At a pause point:

\lstinputlisting[language=rexx,label=pausepoint.rexx,caption=pausepoint.rexx]{pausepoint.rexx}

@@

I/O can be redirected when submitting commands to an external
environment. The submitted command's input stream can be taken from an
existing stream or from a set of compound variables with a common stem.
In the latter case (that is, when a stem is specified as the source for
the commands input stream) whole number tails are used to order input
for presentation to the submitted command. \texttt{Stem.0} must contain
a whole number indicating the number of compound variables to be
presented, and \texttt{stem.\ 1} through \texttt{stem.n} (where
\texttt{n=stem.0}) are the compound variables to be presented to the
submitted command.

@@

The redirections specified on the ADDRESS instruction may not be
possible. If the configuration is aware that the command processor named
does not perform I/O in a manner compatible with the request, the value
of \texttt{\#Env\_Type.} may be set to
\texttt{\textquotesingle{}UNUSED\textquotesingle{}} as an alternative to
\texttt{\textquotesingle{}STEM\textquotesingle{}} and
\texttt{\textquotesingle{}STREAM\textquotesingle{}} where those values
are assigned in the following code.

In the following code the particular use of
\texttt{\#Contains(address,\ expression)} refers to an expression
immediately contained in the \emph{address}.

\lstinputlisting[language=rexx,label=addressinstruct.rexx,caption=addressinstruct.rexx]{addressinstruct.rexx}

@@

The ARG instruction is a shorter form of the equivalent instruction:\\
\texttt{PARSE\ UPPER\ ARG}\emph{\texttt{template\ list}}

@@

Assignment can occur as the result of executing a clause containing an
assignment (see nnn and nnn), or as a result of executing the VALUE
built-in function, or as part of the execution of a PARSE instruction.
Assignment involves an \emph{expression} and a \emph{VAR\_SYMBOL}. The
value of the \emph{expression} is determined; see nnn.

If the \emph{VAR\_SYMBOL} does not contain a period, or contains only
one period as its last character, the value is associated with the
\emph{VAR\_SYMBOL}:

\lstinputlisting[language=rexx,label=assignmentvarpoolzero.rexx,caption=assignmentvarpoolzero.rexx]{assignmentvarpoolzero.rexx}

Otherwise, a name is derived, see nnn. The value is associated with the
derived name:

\lstinputlisting[language=rexx,label=assignmentvarpoolone.rexx,caption=assignmentvarpoolone.rexx]{assignmentvarpoolone.rexx}

@@

The CALL instruction is used to invoke a routine, or is used to control
trapping of conditions.

lf a \emph{vref} is specified that value is the name of the routine to
invoke:

\lstinputlisting[language=rexx,label=callcontains.rexx,caption=callcontains.rexx]{callcontains.rexx}

If a \emph{taken\_constant} is specified, that name is used.

\lstinputlisting[language=rexx,label=calltakenconstant.rexx,caption=calltakenconstant.rexx]{calltakenconstant.rexx}

The name is used to invoke a routine, see nnn. If that routine does not
return a result the RESULT and .RESULT variables become uninitialized:

\lstinputlisting[language=rexx,label=callvardrop.rexx,caption=callvardrop.rexx]{callvardrop.rexx}

@@

If the routine returns a result and the trace setting is
\texttt{\textquotesingle{}R\textquotesingle{}} or
\texttt{\textquotesingle{}I\textquotesingle{}} then a trace with that
result and a tag
\texttt{\textquotesingle{}\textgreater{}\textgreater{}\textgreater{}\textquotesingle{}}
shall be produced, associated with the call instruction.

If a \emph{callon\_spec} is specified:

\lstinputlisting[language=rexx,label=callcallonspec.rexx,caption=callcallonspec.rexx]{callcallonspec.rexx}

@@

For a definition of the syntax of a command, see nnn.

A command that is not part of an ADDRESS instruction is processed in the
ACTIVE environment.

\lstinputlisting[language=rexx,label=commandtoconfifuration.rexx,caption=commandtoconfifuration.rexx]{commandtoconfifuration.rexx}

\texttt{CommandIssue} is also used to describe the ADDRESS instruction:

\lstinputlisting[language=rexx,label=commandIssue.rexx,caption=commandIssue.rexx]{commandIssue.rexx}

@@

The DO instructions is used to group instructions together and
optionally to execute them repeatedly.

Executing a \emph{do\_simple} has the same effect as executing a
\emph{nop}, except in its trace output. Executing the \emph{do\_ending}
associated with a \emph{do\_simple} has the same effect as executing a
\emph{nop}, except in its trace output.

A \emph{do\_instruction} that does not contain a \emph{do\_simple} is
equivalent, except for trace output, to a sequence of instructions in
the following order.

\lstinputlisting[language=rexx,label=doinstruction.rexx,caption=doinstruction.rexx]{doinstruction.rexx}

The following three assignments are made in the order in which
\texttt{\textquotesingle{}TO\textquotesingle{}},
\texttt{\textquotesingle{}BY\textquotesingle{}} and
\texttt{\textquotesingle{}FOR\textquotesingle{}} appear in
\emph{docount}; see nnn.

\lstinputlisting[language=rexx,label=doassignments.rexx,caption=doassignments.rexx]{doassignments.rexx}

@@

When clauses are being traced by \texttt{\#TraceSource}, due to
\texttt{pos(\#Tracing.\#Level,\ \textquotesingle{}AIR\textquotesingle{})\ \textgreater{}\ 0},
the DO instruction shall be traced when it is encountered and again each
time the \texttt{IterateLabel} (see nnn) is encountered. The END
instruction shall be traced when the \texttt{TraceOfEnd} label is
encountered.

When expressions or intermediates are being traced they shall be traced
in the order specified by nnn. Hence, in the absence of conditions
arising, those executed prior to the first execution of
\texttt{OnceLabel} shall be shown once per execution of the DO
instruction; others shall be shown depending on the outcome of the
tests.

The code in the DO description:

\lstinputlisting[language=rexx,label=dolooptracing.rexx,caption=dolooptracing.rexx]{dolooptracing.rexx}

represents updating the control variable of the loop. That assignment is
subject to tracing, and other expressions involving state variables are
not. When tracing intermediates, the BY value will have a tag of
\texttt{\textquotesingle{}\textgreater{}+\textgreater{}\textquotesingle{}}.

@@

The words of the \emph{variable\_list} are processed from left to right.

A word which is a VAR\_SYMBOL, not contained in parentheses, specifies a
variable to be dropped. If VAR\_SYMBOL does not contain a period, or has
only a single period as its last character, the variable associated with
VAR\_SYMBOL by the variable pool is dropped:

\lstinputlisting[language=rexx,label=drop.rexx,caption=drop.rexx]{drop.rexx}

@@

\lstinputlisting[language=rexx,label=drop2.rexx,caption=drop2.rexx]{drop2.rexx}

If the word of the \emph{variable\_list} is a VAR\_SYMBOL enclosed in
parentheses then the value of the VAR\_SYMBOL is processed. The value is
considered in uppercase:

\lstinputlisting[language=rexx,label=dropconfigupper.rexx,caption=dropconfigupper.rexx]{dropconfigupper.rexx}

@@

If the word does not have the syntax of VAR\_SYMBOL a condition is
raised:

\lstinputlisting[language=rexx,label=dropnotsymbolcondition.rexx,caption=dropnotsymbolcondition.rexx]{dropnotsymbolcondition.rexx}

Otherwise the VAR\_SYMBOL indicated by the word is dropped, as if that
VAR\_SYMBOL were a word of the \emph{variable\_list}.

@@

Any \emph{expression} is evaluated:

\lstinputlisting[language=rexx,label=exitevaluate.rexx,caption=exitevaluate.rexx]{exitevaluate.rexx}

The opportunity is provided for a final trap.

\lstinputlisting[language=rexx,label=finaltrap.rexx,caption=finaltrap.rexx]{finaltrap.rexx}

The processing of the program is complete. See nnn for what
\texttt{API\_Start} returns as the result.

@@

\emph{We need a check that this starts method; similarities with
PROCEDURE.}

@@

It is used at the start of a method, after method initialization, to
make variables in the receiver's pool accessible:

\lstinputlisting[language=rexx,label=allowexpose.rexx,caption=allowexpose.rexx]{allowexpose.rexx}

The words of the \emph{variable\_list} are processed from left to right.

A word which is a VAR\_SYMBOL, not contained in parentheses, specifies a
variable to be made accessible. If VAR\_SYMBOL does not contain a
period, or has only a single period as its last character, the variable
associated with VAR\_SYMBOL by the variable pool (as a non-tailed name)
is given the attribute `exposed'.
